The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-02-17 analysis/paper/paper.qmd
Add script for tf-idf statistics
The tf-idf is used to show high-frequency words that are document-specific in a larger collection of documents.

>>>>

```{r}
#| label: tf-idf

emu_theses_words

emu_theses_words_count <- emu_theses_words |> 
  count(title, word, sort = TRUE) |> 
  ungroup()

emu_theses_words_total <- emu_theses_words_count |> 
  group_by(title) |> 
  summarise(total = sum(n))

emu_theses_words_count <- left_join(emu_theses_words_count, emu_theses_words_total)

# emu_theses_words_count |> 
#   group_by(title) |> 
#   summarise(total = unique(tot

# Preview term frequency distributions in a subset of theses
ggplot(emu_theses_words_count |> 
         filter(title == unique(title)[1:6]),
       aes(n/total, fill = title)) +
  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
  xlim(NA, 0.0009) +
  facet_wrap(~ title, ncol = 3, scales = "free_y")

# Examine Zipf's law for the EMU theses
## Calculate rank and term frequency
freq_by_rank <- emu_theses_words_count |> 
  group_by(title) |> 
  mutate(rank = row_number(),
         `term frequency` = n/total)

## Visualise Zipf's law
freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = title)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  scale_x_log10() +
  scale_y_log10()

rank_subset <- freq_by_rank |> 
  filter(rank < 500,
         rank > 10)

lm(log10(`term frequency`) ~ log10(rank), data =  rank_subset)

freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = title)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
  scale_x_log10() +
  scale_y_log10()

## Calculate tf-idf
emu_theses_words_count <- emu_theses_words_count |> 
  bind_tf_idf(word, title, n)
emu_theses_words_count

emu_theses_words_count |> 
  select(-total) |> 
  arrange(desc(tf_idf))

## Visualise the top n words with the highest tf-idf values for each of a subset of theses
emu_theses_words_count |> 
  filter(title == unique(title)[1:6]) |> 
  arrange(desc(tf_idf)) |> 
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
  group_by(title) |> 
  top_n(15) |> 
  ungroup() |> 
  ggplot(aes(word, tf_idf, fill = title)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~ title,  ncol = 3, scales = "free") +
  coord_flip()
```

The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words. As such, it shows words that are the most important to one document in a collection of documents [@silge2017].
